---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# smcepi

<!-- badges: start -->
<!-- badges: end -->

The goal of smcepi is to provide functions that make data analysis easier for San Mateo County epidemiologists.  

## Installation

You can install the development version of smcepi from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("San-Mateo-County-Health-Epidemiology/smcepi")
```

## Using `theme_gg_smc()`

`theme_gg_smc()` will format your `ggplot2` charts according to OEE style guidelines. In order to use the county's fonts, you'll need to run `load_smc_fonts()` once per session so that the fonts are available for use in the `theme_gg_smc()` function. 

```{r example, out.width='60%', fig.align='center'}
library(smcepi)

load_smc_fonts()

iris %>% 
  ggplot(aes(x = Species, 
             y = Sepal.Length)) +
  geom_point(
    position = position_jitter(w = 1/6, h = 0),
    size = 4, 
    color = "#006cb6", 
    alpha = 2/3
  ) +
  scale_y_continuous(
    limits = c(4, 8),
    breaks = seq(4, 8, 2)
  ) + 
  labs(title = "Iris Jitter Plot") + 
  theme_gg_smc()

```


