---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# smcepi

<!-- badges: start -->

<!-- badges: end -->

The goal of smcepi is to provide functions that make data analysis easier for San Mateo County epidemiologists.

## Installation

You can install the development version of smcepi from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("San-Mateo-County-Health-Epidemiology/smcepi")
```

## Using formatting themes

### `theme_gg_smc()`

`theme_gg_smc()` will format your `ggplot2` charts according to OEE style guidelines. Embedded within the `theme_gg_smc()` is the `load_smc_fonts()` function that will download and/or enable the fonts from the [extrafont](https://cran.r-project.org/web/packages/extrafont/extrafont.pdf) package. If you haven't already installed the extrafont fonts on your computer, the `theme_gg_smc()` and `load_smc_fonts()` functions will download them the first time you run either of those functions. 

```{r theme_gg_smc-example, out.width='60%', fig.align='center'}
library(smcepi)

iris %>% 
  ggplot(aes(x = Species, 
             y = Sepal.Length)) +
  geom_point(
    position = position_jitter(w = 1/6, h = 0),
    size = 4, 
    color = "#006cb6", 
    alpha = 2/3
  ) +
  scale_y_continuous(
    limits = c(4, 8),
    breaks = seq(4, 8, 2)
  ) + 
  labs(title = "Iris Jitter Plot") + 
  theme_gg_smc(plot_lines = "vertical")

```

### `theme_ft_smc()`

`theme_ft_smc()` will format your `flextable` charts according to OEE style guidelines. Like in the `smc_gg_theme()` function, the `load_smc_fonts()` is embedded within the `theme_ft_smc()` function. 

```{r theme_ft_smc-example, out.width='60%', fig.align='center'}
library(smcepi)

iris %>%
  dplyr::slice(1:5) %>%
  dplyr::select(Species, everything()) %>%
  flextable() %>%
  footnote(i = 1, j = 1, value = as_paragraph("footer test")) %>%
  theme_ft_smc()

```

## Selectively coloring text

### `gg_color_title()`

`gg_color_title()` allows you to specify the color for specific words or phrases in your `ggplot2` title, subtitle or caption. 

In order for the `gg_color_title()` to render correctly in your chart, you should use the `ggtext::element_markdown()` in your `ggplot2::theme()` instead of using the standard `element_markdown()`. The `theme_gg_smc()` function uses `element_markdown()`.

```{r gg_color_title-example, out.width='60%', fig.align='center'}
library(ggtext)

brown_color <- "#873600"
blue_color <- "#006cb6"

title <- gg_color_title("Brown and blue eyes are most common",
                       c("Brown", "blue"), c(brown_color, blue_color))
HairEyeColor %>%
  data.frame() %>%
  dplyr::group_by(Eye) %>%
  dplyr:: summarize(freq = sum(Freq),
                    .groups = "keep") %>%
  dplyr:: ungroup() %>%
  ggplot2::ggplot(aes(x = Eye,
                      y = freq,
                      fill = Eye)) +
  scale_fill_manual(values = c("Brown" = brown_color,
                               "Blue" = blue_color,
                               "Hazel" = "#D5D8DC",
                               "Green" = "#D5D8DC")) + 
  geom_bar(position = "dodge",
           stat = "identity") +
  labs(title = title) +
  theme_gg_smc() +
  ggplot2::theme(plot.title = ggtext::element_markdown(size = 16, hjust = 0, face = "bold"))
```
