---
title: "Azure Helpers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Azure Helpers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The Azure helper functions are meant to help people who process data primarily in R but store data in Azure. The functions are wrapper functions that lean heavily on the [DBI](https://dbi.r-dbi.org/) package. 

## `connect_azure()`

### Basic usage

This is a wrapper around the `DBI::dbConnect()` function. The only required input is a `.yml` file that contains credentials for connecting to the database. 

The function expects a `.yml` file that includes values for: `server`, `driver`, `authenticator`, `database`, `uid_suffix` and `port`. Here's an example `.yml` file template:

```{r eval=FALSE}
default:
  db:
    server: 'server-name'
    driver: 'ODBC Driver 18 for SQL Server'
    authenticator: 'ActivedirectoryPassword'
    database: 'database-name'
    uid_suffix: '@organization.org'
    port: port-number
```

When you pass the `.yml` file through the function, you'll be prompted to enter your Windows password and then will be connected to the Azure database you've specified in the `.yml` file.

```{r eval=FALSE}
azure_con <- connect_azure(creds_file = "*.yml")
```

### Additional options
### `creds_position`

If you have multiple sets of credentials stored in your `.yml` file, you can select which set you'd like to use by specifying the position of the credentials.

Let's say your `.yml` file looks like this:

```{r eval=FALSE}
default:
  db_1:
    server: 'server-name'
    driver: 'ODBC Driver 18 for SQL Server'
    authenticator: 'ActivedirectoryPassword'
    database: 'database-name'
    uid_suffix: '@organization.org'
    port: port-number
  
  db_2:
    server: 'server-name'
    driver: 'ODBC Driver 18 for SQL Server'
    authenticator: 'ActivedirectoryPassword'
    database: 'database-name'
    uid_suffix: '@organization.org'
    port: port-number
```

If you want to use the second set of credentials (`db_2`) instead of the first set (`db_1`) you should set `creds_position = 2`:

```{r eval=FALSE}
azure_con <- connect_azure(creds_file = "*.yml",
                           creds_position = 2)
```

#### `pwd`

By default, `connect_azure()` allows you to manually enter your password when you run the function. If you prefer to use a stored password, you can override the default using the `pwd` argument.

```{r eval=FALSE}
azure_con <- connect_azure(creds_file = "*.yml",
                           pwd = stored_password)
```

## `azure_import_loop()`

This function creates a table in Azure from R. It is a wrapper function for the `DBI::dbWriteTable()` and the `DBI::dbAppendTable()` functions. First, this function breaks the dataset to be imported into smaller groups and then it loops through the groups. The first group is used to create the table and subsequent groups are appended to the table.

This function is helpful for writing large datasets from R to Azure. Small datasets can be sent from R to Azure just using the `DBI::dbWriteTable()` function.

```{r eval=FALSE}
azure_import_loop(azure_con = azure_con,
                  data = data,
                  group_size = 500,
                  table_name = "new_azure_table_name",
                  field_types = varchar_max)
```

Arugments:

* `azure_con`: A DBI Connection object as returned by `dbConnect()`. You can also create this using the `smcepi::connect_azure()` function. See help text for `DBI::dbWriteTable()` for technical details. 
* `data`: A data.frame object you plan to import
* `group_size`: An integer that indicates the size of the import groups. Ex: a group_size of 50 would split the dataset into groups of 50 records and would then import each group at a time
* `table_name`: A string to specify the name of the table in Azure
* `field_types`: Optional parameter to pass a list of varchar(`n`) lengths for the Azure table. Can be created with the `smcepi::varchar_max()` function

## `varchar_max()`

Use this to generate a list of varchar(`n`) specifications to be used in an R to Azure import with the `DBI:dbWriteTable()` or `smcepi::azure_import_loop()` functions. 

This list specifies a `varchar(n)` value for each character variable based on the maximum string length in that variable. `n` is 255 for variables with maximum string lengths of 255 or less and `n` is the maximum string length for variables with maximum string lengths longer than 255.

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

