---
title: "Data Management"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data File Management}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(smcepi)
```

### remove_right_empty_cols()

This function removes all empty columns on the right of a data set. This function is useful when you import a file and R adds empty columns to the right of the actual data. This function won't remove empty columns on the left or in the middle of the data set. 

#### it will remove empty columns on the right...

This data has two empty columns on the right:

```{r}
data <- data.frame(
  col1 = sample(1:100, 100, replace = T),
  col2 = NA,
  col3 = runif(100),
  col4 = sample(c("blue", "green", "yellow", "orange"), 100, replace = T),
  col5 = NA,
  col6 = NA
)
head(data)
```

After passing it through `remove_right_empty_cols()`, the two empty columns on the right (`col5` and `col6` have been dropped):

```{r}
head(remove_right_empty_cols(data))
```

#### ... and only on the right.

If you accidentally run this function on a data set that doesn't have any empty columns on the right, the data set will not change. The rightmost variable in this data set has information in it:

```{r}
data <- data.frame(
  col1 = sample(1:100, 100, replace = T),
  col2 = NA,
  col3 = runif(100),
  col4 = NA,
  col5 = NA,
  col6 = sample(c("blue", "green", "yellow", "orange"), 100, replace = T)
)

head(data)
```

In this example `col6` has data in it, so no variables are dropped from the data set. 

```{r}
head(remove_right_empty_cols(data))
```
