---
title: "Life Expectancy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Life Expectancy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(smcepi)
```

Public Health England (PHE) has an excel file that will calculate life expectancy using an abridged life table. The `make_life_table()` function in this package was written to exactly replicate the formulas and methods used in that excel file. The PHE excel file can be found  [here](https://fingertips.phe.org.uk/documents/phe%20life%20expectancy%20calculator.xlsm). 

The `make_life_table()` function is designed to work with grouped data so that you can calculate life tables for multiple groups at once. It also allows for the use of ACS age groups in addition to the age groups included in the PHE excel calculator. 

* PHE age groups: 0, 1-4, 5-9, 10-14 ... 80-84, 85-89, 90+  
* ACS age groups: 0-4, 5-9, 10-14 ... 80-84, 85+

## make_life_table()
### Basic usage 

The output from the `make_life_table()` function has every column that is included in the PHE excel calculator. This example includes the same death and population numbers that are in the PHE excel calculator:

```{r}
data1 <- data.frame(
  age_cat = c("0", "1-4", "5-9", "10-14", "15-19","20-24", "25-29", "30-34", "35-39", "40-44",
  "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80-84", "85-89", "90+"),
  deaths = c(206, 37, 23, 23, 105, 162, 268, 314, 413, 584, 954,
            1359, 1912, 2824, 4507, 5851, 7117, 8192, 7745, 6442),
  population = c(50698, 215400, 280458, 258105, 282062, 329060, 306097, 274544, 260415,
           267450, 311314, 324311, 296825, 271339, 284608, 228062, 162785, 111263, 58987, 26016))

life_table1 <- make_life_table(data1)
str(life_table1)
```

### Additional options
#### Multiple groups
One benefit to using the `make_life_table()` function instead of the excel file is that you can create full life tables for multiple groups at once. You just specify the grouping variables with the `grouping_vars` argument. 

Here we have a version of `data1` but with an additional set of data that were simulated based on the data entered into the PHE excel file. 

```{r}
data2 <- data.frame(
  group = c(rep("phe", 20), rep("simulated", 20)),
  age_cat = rep(c("0", "1-4", "5-9", "10-14", "15-19","20-24", "25-29", "30-34", "35-39", "40-44",
                 "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80-84", "85-89", "90+"), 2),
  deaths = c(206, 37, 23, 23, 105, 162, 268, 314, 413, 584, 954,
               1359, 1912, 2824, 4507, 5851, 7117, 8192, 7745, 6442,
               232, 30, 41, 22, 194, 168, 315, 313, 406, 643, 963,
               1446, 1979, 2814, 4587, 5874, 7111, 8221, 7825, 6540),
  population = c(50698, 215400, 280458, 258105, 282062, 329060, 306097, 274544, 260415,
                       267450, 311314, 324311, 296825, 271339, 284608, 228062, 162785, 111263, 58987, 26016,
                       51578, 215512, 279462, 257256, 282348, 329111, 306514, 274397, 259847, 267045,
                       311791, 323739, 297453, 271344, 285047, 227655, 162922, 110554, 58886, 26243))
```

Then you can create life tables for each group by specifying that `grouping_vars = "group"`. The `grouping_vars` argument can take a list of variables if you want to group by year and geography (for example). 

```{r}
life_table2 <- make_life_table(data2,
                               grouping_vars = "group")
str(life_table2)
```
#### Renaming columns

The function expects a `data.frame` with a minimum of three columns:

- age categories (`age_cat`): this should be a character variable with the start age and end age separated by a `-`. Ex: `c("0", "1-4", "5-9", "10-14", ..."90+")`.
- death counts (`deaths`): this should be a numeric variable with the number of deaths that occurred in each age group for each interval
- population years (`population`): this should be a numeric variable with the number of population years in each group for each interval

If your variables aren't named `"age_cat"`, `"deaths"` and `"population"`, you can pass your variable names to the function using the `age_cat_var`, `deaths_var` and `population_var` arguments. For example:

```{r}
data2 <- data.frame(
  ages = c("0", "1-4", "5-9", "10-14", "15-19","20-24", "25-29", "30-34", "35-39", "40-44",
  "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80-84", "85-89", "90+"),
  death_count = c(206, 37, 23, 23, 105, 162, 268, 314, 413, 584, 954,
            1359, 1912, 2824, 4507, 5851, 7117, 8192, 7745, 6442),
  population_years = c(50698, 215400, 280458, 258105, 282062, 329060, 306097, 274544, 260415,
           267450, 311314, 324311, 296825, 271339, 284608, 228062, 162785, 111263, 58987, 26016))

life_table1 <- make_life_table(data2,
                               age_cat_var = "ages",
                               deaths_var = "death_count",
                               population_var = "population_years")
str(life_table1)
```

## get_le()

### Basic usage

Typically you just want the estimated life expectancy at age 0 and you don't need the entire the life table. You can use the `get_le()` function on the output of the `make_life_table()` function for this. Note the `get_le()` function was written explicitly to work with the output of the `make_life_table()` function. 

```{r}
get_le(life_table1)
```

### Additional options

#### Multiple groups

`get_le()` also works for grouped data. Like with the `make_life_table()` function, you should use the `grouping_vars` argument to specify the groups: 

```{r}
get_le(life_table2, grouping_vars = "group")
```

#### Excluding confidence intervals

95% confidence intervals are included in the output of `get_le()` by default. You can elect to just get the estimate by setting `include_ci` to `FALSE`.

```{r}
get_le(life_table1, include_ci = F)
```

#### Ages other than 0 

If you're interested in the life expectancy at an age group other than 0, you can specify the age with the `start_age` argument. The start age you specify should be numeric and must be the start of an age group in the data passed through `make_life_table()`.

```{r}
get_le(life_table1, start_age = 30)
```
