---
title: "Relative Standard Error"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Relative Standard Error}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")

library(smcepi)
```

This function calculates the relative standard error for sample and survey data.

## Data structure-Sample

This data includes a standard deviation, a sample size, and a method ("sample" in this case) to determine the appropriate calculation for the data.

```{r}
  rse_df_sample <- data.frame(
    sd = 25,
    n = 1000,
    method = "sample")
```

## Application of the function

To use the function, supply the standard deviation, sample size, and method to the `prop_rse` function.

```{r}
prop_rse(rse_df_sample$sd, rse_df_sample$n, rse_df_sample$method)
```

## Data structure-Survey

This data includes a standard deviation, a sample size, and a method ("survey" in this case) to determine the appropriate calculation for the data.

```{r}
  rse_df_survey <- data.frame(
    sd = 25,
    n = 1000,
    method = "survey")
```

## Application of the function

To use the function, supply the standard deviation, sample size, and method to the `prop_rse` function.

```{r}
prop_rse(rse_df_survey$sd, rse_df_survey$n, rse_df_survey$method)
```
