---
title: "San Mateo Geography Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{San Mateo Geography Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The San Mateo County geography functions are designed to help clean and categorize geographic variables consistently.  

## `smc_city_clean()`

### Basic usage

This function uses regular expressions to clean up San Mateo County city names. Non-San Mateo city names will return an `NA` value. By default, the function assumes the variable you want to clean is called `city` and will save the cleaned results in a variable called `city_clean`.

```{r eval=FALSE}
data <- data.frame(
  city = c("Burligame", "Fost City", "San Mato", "Daily Cit", "S S Francisco", "South San Fransico", "SoSan Franc", "San Francisco")
  ) 

data %>%
  smc_city_clean()

```

### Additional options

If your input column is not called `city` you should pass the name of your city variable in the `city_col` argument. You also have the option of specifying the name of the variable for the cleaned cities. By default it is `city_clean`. 

```{r eval=FALSE}
data <- data.frame(
  city_dirty = c("Burligame", "Fost City", "San Mato", "Daily Cit", "S S Francisco", "South San Fransico", "SoSan Franc", "San Francisco")
  ) 

data %>%
  smc_city_clean(city_col = "city_dirty",
                 new_col = "smc_city")

```


## `smc_zip_region_sort()`

### Basic usage

This function categorizes zip codes into county regions. The expected input is a variable in a data frame and it will return a second variable with the zip region. The region options are: North, Mid, South, Coastside and "Not a residential zip" (for PO Boxes). 

```{r eval=FALSE}
data <- data.frame(
  zip = c("94015", "94403", "94303", "94019", "94128", "94110")
  )

data %>%
  smc_zip_region_sort()

```

### Additional options

By default, the function expects the variable of zip codes to be called `zip` and it will save the zip regions in a variable called `zip_region`. However, if your input column is not called `zip` you can override the default with the `zip_col` argument. You can also change the outputted variable name with the `region_col` argument. 

```{r eval=FALSE}
data <- data.frame(
  smc_zip = c("94015", "94403", "94303", "94019", "94128", "94110")
  )

data %>%
  smc_zip_region_sort(zip_col = "smc_zip",
                      region_col = "smc_zip_region")

```

